---
title: "Analysis of the 1933 German Election during the Weimar Republic"
output: html_document
---

 Who voted for the Nazis?  Researchers attempted to answer this
question by analyzing aggregate election data from the 1933 German
election during the Weimar Republic.

This exercise is based on the following article: King, Gary, Ori Rosen, Martin Tanner, Alexander F. Wagner. 2008.  "[Ordinary Economic Voting Behavior in the Extraordinary Election of Adolf Hitler.](http://dx.doi.org/10.1017/S0022050708000788)" *Journal of Economic History* 68(4): 951-996.  

We analyze a simplified version of the election outcome data, which records, for each precinct, the number of eligible voters as well as the number of votes for the Nazi
party.  In addition, the data set contains the aggregate occupation
statistics for each precinct. The table below presents the
variable names and descriptions of the data file
`nazis.csv`.  Each observation represents a German precinct.

------------------------------------------------------------------------------
 Name                Description
 ------------------- ---------------------------------------------------------
 `shareself`         Proportion of self-employed potential voters
 
 `shareblue`         Proportion of blue-collar potential voters
 
 `sharewhite`        Proportion of white-collar potential voters
 
 `sharedomestic`     Proportion of domestically employed potential voters
 
 `shareunemployed`   Proportion of unemployed potential voters
 
 `nvoter`            Number of eligible voters
 
 `nazivote`          Number of votes for Nazis
------------------------------------------------------------------------------
 
The goal of analysis is to investigate which types of voters (based on
their occupation category) cast ballots for the Nazis.  One hypothesis
says that the Nazis received much support from blue-collar workers.
Since the data do not directly tell us how many blue-collar workers
voted for the Nazis, we must infer this information using a
statistical analysis with certain assumptions.  Such an analysis where
researchers try to infer individual behaviors from aggregate data is
called *ecological inference*.

To think about ecological inference more carefully in this context,
consider the following simplified table for each precinct $i$:
 
 Vote-Choice                  Blue-Collar   Non-blue-collar   
 ---------------------------- ------------- ----------------- -----------
 Nazis                        $W_{i1}$      $W_{i2}$          $Y_i$     
 Other parties or abstention  $1-W_{i1}$    $1-W_{i2}$        $1-Y_i$  
                              $X_i$         $1-X_i$
 ------------------------------------------------------------------------
                              
The data at hand only tells us the proportion of blue-collar voters
$X_i$ and the vote share for the Nazis $Y_i$ in each precinct, but we
would like to know the Nazi vote share among the blue-collar voters
$W_{i1}$ and among the non-blue-collar voters $W_{i2}$.  Then, there
is a deterministic relationship between $X$, $Y$, and $W_1, W_2$.
Indeed, for each precinct $i$, we can express the overall Nazis' vote
share as the weighted average of the Nazis' vote share of each
occupation,

$$
  Y_i = X_i W_{i1} + (1-X_i) W_{i2}
$$

## Question 1

We exploit the linear relationship between the Nazis' vote share
  $Y_i$ and the proportion of blue-collar voters $X_i$ given in
  the equation above by regressing the former on
  the latter.  That is, fit the following linear regression model,
  
  $$
    E(Y_i | X_i) = \alpha + \beta X_i
  $$
  Compute the estimated slope coefficient, its standard error, and the
  95% confidence interval.  Give a substantive interpretation of each
  quantity.
  
## Answer 1

```{r}
nazis <- read.csv("data/nazis.csv")
nazis$sharenazis <- nazis$nazivote / nazis$nvoter # Y_i as vote share
simplefit <- lm(sharenazis ~ shareblue, data = nazis)
summary(simplefit)
confint(simplefit)
```

The simple regression indicates that as the percentage of blue-collar
voters increases, so too does the Nazi vote share. However, the
standard error is about 0.05, which is fairly large.

A better understanding of this slope coefficient requires the readers
to directly connect the two equations above.
To do this, take the conditional expectation of the former given
$X_i$:

$$
  E(Y_i | X_i) = E(W_{i1} | X_i) X_i + E(W_{i2} | X_i) (1-X_i) \\
  = \underbrace{E(W_{i2} | X_i)}_{= \ \ \ \alpha} + \underbrace{E(W_{i1} - W_{i2} | X_i)}_{= \ \ \ \beta} X_i
$$

Thus, the model assumes that the Nazi expected vote share among
blue-collar voters (as well as that among non-blue-collar voters) does
not depend on the proportion of blue-collar voters in the precincts.
That is, $E(W_{i1} | X_i)= E(W_{i1})$ and
$E(W_{i2} | X_i) = E(W_{i2})$ are assumed to be constant and do not
depend on $X_i$.  Under this model, the slope coefficient $\beta$
represents the difference in the Nazis' expected vote share among
blue-collar voters and among other voters.  The result indicates that
blue-collar voters are more likely to cast a ballot for the Nazis on
average, but the difference is not statistically significant at the
95% level.

## Question 2

Based on the fitted regression model from the previous question,
  predict the average Nazi vote share $Y_i$ given various proportions
  of blue-collar voters $X_i$.  Specifically, plot the predicted value
  of $Y_i$ (the vertical axis) against various values of $X_i$ within
  its observed range (the horizontal axis) as a solid line.  Add 95\%
  confidence intervals as dashed lines.  Give a substantive
  interpretation of the plot.
  
## Answer 2

```{r}
blue.range <- seq(from = 0, to = 1, by = 0.01)
blue.pred <- predict(simplefit, interval = "confidence", 
                     newdata = data.frame(shareblue = blue.range))
plot(blue.range, blue.pred[, "fit"], col = "red",
     xlim = c(0, 1), ylim = c(0.3, 0.6), type = "l", 
     xlab = "Blue-Collar Share", ylab = "Nazi Vote Share",
     main = "Predicted Blue-Collar Support") # predicted
lines(blue.range, blue.pred[, "lwr"], lty = "dashed", col = "blue")
lines(blue.range, blue.pred[, "upr"], 
      lty = "dashed", col = "blue") # confidence intervals
```

The graph indicates that as the blue-collar share goes up, so too does
the Nazi vote-share, although the wide confidence intervals suggest
that the evidence for this relationship is rather weak.

## Question 3

Fit the following alternative linear regression model,

  $$
    E(Y_i | X_i) = \alpha^\ast X_i + (1-X_i) \beta^\ast
  $$
  Note that this model does not have an intercept.  How should one
  interpret $\alpha^\ast$ and $\beta^\ast$?  How are these parameters
  related to the linear regression model given in Question 1?
  
## Answer 3

Comparing this equation  with the equations above shows that under this alternative linear
regression model we have $E(W_{1i} \mid X_i) = \alpha^\ast$ and
$E(W_{2i} | X_i)=\beta^\ast$.  Thus, the relationship between the
two linear models is characterized by $\alpha = \beta^\ast$ and
$\beta = \beta^\ast - \alpha^\ast$.  Both models are based on the
assumption that the average proportion of voters voting for Nazis
among each occupational group does not depend on the occupational
composition of each precinct, i.e., $E(W_{ij} | X_i)= E(W_{ij})$
for $j=1,2$.

```{r}
nazis$shareothers <- 1 - nazis$shareblue
altfit <- lm(sharenazis ~ -1 + shareblue + shareothers, 
             data = nazis)
summary(altfit)
confint(altfit)
```

The result implies that `r round(coef(altfit)["shareblue"], 3)*100`% of blue-collar voters were estimated to have voted for the Nazi party.  The 95% confidence interval is given by [`r round(confint(altfit)["shareblue",], 3)*100`].

## Question 4

Fit a linear regression model where the overall Nazi vote share
  is regressed on the proportion of each occupation.  The model should
  contain no intercept and five predictors, each representing the
  proportion of a certain occupation type.  Interpret the estimate of
  each coefficient and its 95% confidence interval.  What assumption
  is necessary to permit your interpretation?
  
## Answer 4

```{r}
fit <- lm(sharenazis ~ -1 + shareself + shareblue + sharewhite + 
              sharedomestic + shareunemployed, data = nazis)
summary(fit) # coefficients
confint(fit) # 95% confidence intervals
```

It seems that support for the Nazis was highest among self-employed
voters, whereas very few domestic employees and unemployed individuals
supported Nazis. However, the coefficients do not quite make sense, as
they imply that if a precinct was 100% self-employed, then 111% of
voters would vote Nazi, and if a precinct was 100% unemployed, then
negative 3% of voters would vote Nazi. This could be a sign of a
possible violation of the assumption.  It could also be the case that
sampling variability leads to the estimates being outside of the
range.

As before, we make the assumptions that the average proportion of 
Nazi voters does not depend on the occupational composition of each 
precinct, and that there is exogeneity, i.e. the conditional 
expectation of the error term given the explanatory variables is the 
same as its unconditional expectation.

## Question 5

Finally, we consider a model-free approach to ecological
  inference.  That is, we ask how much we can learn from the data
  alone without making an additional modeling assumption. For each
  precinct, obtain the smallest value that is logically possible for
  $W_{i1}$ by considering the scenario in which all non-blue-collar
  voters in precinct $i$ vote for the Nazis.  Express this value as a
  function of $X_i$ and $Y_i$.  Similarly, what is the largest
  possible value for $W_{i1}$?  Calculate these bounds, keeping in
  mind that the value for $W_{i1}$ cannot be negative or greater than
  1.  Finally, compute the bounds for the nation-wide proportion of
  blue-collar voters who voted for the Nazis (i.e., combining the
  blue-collar voters from all precincts by computing their weighted
  average based on the number of blue-collar voters).  Give a brief
  substantive interpretation of the results.
  
## Answer 5

To compute the bounds, we solve for $W_{i1}$ within each precinct,
yielding $W_{i1} = \frac{(Y_i - (1 - X_i) W_{i2})}{X_i}$.  The minimum
and maximum value of $W_{i1}$ are obtained by setting $W_{i2} = 1$
(all non-blue collars vote for the Nazis) and $W_{i1} = 0$ (none of
non-blue-collars vote for the Nazis), respectively. We then adjust the
minimum and maximum values of $W_{i1}$ to 0 and 1, respectively, if
those estimates are negative or above 1.  In summary, the bounds are
given by:


$$
  \max(0, \ \frac{Y_i + X_i - 1}{X_i}) \ \le \ W_{i1} \ \le \ \min(1, \frac{Y_i}{X_i})
$$

```{r}
## calculate bounds
blue.min <- (nazis$sharenazis + nazis$shareblue - 1) / nazis$shareblue
blue.min[blue.min < 0] <- 0
blue.max <- nazis$sharenazis / nazis$shareblue
blue.max[blue.max > 1] <- 1
## compute nation-wide proportion bounds
num.blue <- nazis$shareblue * nazis$nvoter # number of blue-collar voters
c(weighted.mean(blue.min, num.blue), 
  weighted.mean(blue.max, num.blue))
```

The no-assumption bounds for $W_{i1}$ are very large, essentially
ranging from 0 to 1, so this method of analysis does not tell us very
much about blue-collar support for the Nazis.  In other words, the
results based on the linear regression model critically rely on the
assumption that the average vote shares among each occupation group
are independent of their relative size within a precinct.
