---
ttitle: "Prediction tidyverse"
author: "Anna Yorozuya"
institute: "University of Tokyo"
date: "June 2022"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
---

# Prediction 1 (June 9)
## Table of Contents
- `stringr` package
- `lubridate` package
- Combining objects
- Today's in-class assignment: `betting-markets`


## `stringr` package
### `stringr` package
- a package in tidyverse for string manipulation
- here we use the function `str_c()`
- for more information, see [here](https://r4ds.had.co.nz/strings.html)

### `str_c` function
- combining multiple objects into a character string
- each object will be separated by commas

### example
```{r include = FALSE}
library(tidyverse)
```
```{r}
a <- 2
b <- 6
print(str_c(a, " times 3 is equal to ", b))
```

## `lubridate` package
### `lubridate` package
- a package in tidyverse for dates-times object
- for more information, see [here](https://r4ds.had.co.nz/dates-and-times.html)

### `ymd` function
- converting strings in the format of `year-month-date` into a `date` object
- when subtracting a `date` object from another, the object will be a `difftime` object.

```{r include = FALSE}
library(lubridate)
```
```{r}
x <- ymd("2008-11-04")
y <- ymd("2008/9/1")
subtraction <- x - y
class(x)
class(subtraction)
```


## Combining objects (data frame, vector, or matrix)
### base R function
- `cbind(x, y)`: combine x and y by columns
- `rbind(x, y)`: combine x and y by rows
- `merge(x, y, by = )`: merging two data frames by common columns or row names

### tidyverse function `bind` (be careful!)
- `bind_cols(x, y)`: combine x and y by columns
- `bind_rows(x, y)`: combine x and y by rows

### tidyverse function `mutate-joins`
- `inner_join(x, y, by = ): includes all rows in x and y
- `left_join (x, y, by = ): includes all rows in x
- `right_join (x, y, by = )`: includes all rows in y
- `full_join (x, y, by = )`: includes all rows in x or y


# Prediction 2 (June 14)
## Table of Contents
- `broom` package
- `modelr` package
- `tidyr` package
- `pivot_longer()` and `pivot_wider()`
- visualizing regression
- Today's in-class assignment: `conditional-cash-transfer`


## `broom` package
### what is `broom`?
- a package in `tidymodels` package
- converting outputs of baseR functions into tidy data
- for more information, see [here](https://broom.tidymodels.org/index.html).

### useful functions
- `tidy()`: summarizes information about model components
- `glance()`: reports information about the entire model
- `augment()`: adds informations about observations to a dataset

### useful functions, when used for `lm()` outputs
- `tidy()`: returns a data frame in which each row is a coefficient
- `glance()`: returns a one-row dataframe summary of the model
- `augment()`: returns the original data with fitted values, residuals, and other observation level stats from the model appended to it.


## `broom` package: example
```{r include = FALSE}
data("face", package = "qss")
face <- mutate(face,
                d.share = d.votes / (d.votes + r.votes),
                r.share = r.votes / (d.votes + r.votes),
                diff.share = d.share - r.share)
library(tidymodels)
```

```{r}
fit <- lm(diff.share ~ d.comp, data = face)
glance(fit)
```

## `broom` package: example
```{r}
tidy(fit)
```

## `broom` package: example
```{r}
augment(fit) %>% head()
```


## `modelr` package
### what is `modelr`?
- a package for helping modelling in tidyverse framework, especially with pipes
- for more information, see [here](https://r4ds.had.co.nz/model-basics.html)

### useful functions
- `add_predictions()`: add the predictions to the original data
- `add_residuals()`: add the residuals to the original data
- `data_grid()`: create a data set containing every unique combination of the specified columns from the old data set.
- `spread_predictions()`: generate two sets of predictions for a new tibble of data

## `modelr` package: example
```{r include = FALSE}
library(modelr)
data("florida", package = "qss")
data("social", package = "qss")
```
```{r}
fit2 <- lm(Buchanan00 ~ Perot96, data = florida)
florida_fit2 <- florida %>%
  add_predictions(fit2) %>%
  add_residuals(fit2)
head(florida_fit2)
```
```{r}
fit <- lm(primary2006 ~ messages, data = social)
unique_messages <- data_grid(social, messages) %>%
  add_predictions(fit)
unique_messages
```


## `tidyr` package
### what is `tidyr`?
- a package in tidyverse helping to tidy data
- for more data, see [here](https://r4ds.had.co.nz/tidy-data.html) 

### useful function
- `crossing()`: produce a new data set with all combinations of the specified variable values

## `tidyr` package: example
```{r include=FALSE}
social.neighbor <- 
  filter(social, 
         messages == "Control" | messages == "Neighbors") %>%
  mutate(age = 2008 - yearofbirth)
```
```{r}
fit.age <- lm(primary2006 ~ age * messages, data = social.neighbor)
ate.age <- tidyr::crossing(age = seq(from = 20, to = 80, by = 20),
         messages = c("Neighbors", "Control")) %>%
  add_predictions(fit.age) %>%
  pivot_wider(names_from = messages,
              values_from = pred) %>%
  mutate(diff = Neighbors - Control)
ate.age
```

## `pivot_longer()` and `pivot_wider()`
### `pivot_longer()`
- increase the number of rows, while decreasing the number of columns
- argument `cols = x`: specify the columns (x) to pivot into longer formats
- argument `names_to`: name the new columns for storing data from the columns specified in the `cols` argument.

### `pivot_wider()`
- increase the number of columns, while decreasing the number of rows
- argument `names_from`: describe which column to get the name of the output column.

## `pivot_longer()` and `pivot_wider()`: example
```{r include = FALSE}
data("women", package = "qss")
library(tidyverse)
```
```{r}
women %>%
  group_by(reserved) %>%
  summarize(irrigation = mean(irrigation),
            water = mean(water)) %>%
  pivot_longer(names_to = "variable", - reserved) %>%
  pivot_wider(names_from = reserved) %>%
  rename("not_reserved" = `0`,
         "reserved" = `1` ) %>%
  mutate(diff = reserved - not_reserved)
```


## visualizing regression 1: `geom_point()` + `geom_abline()`
```{r}
ggplot() +
  geom_point(data = face,
             mapping = aes(x = d.comp, y = diff.share), shape = 1) +
  geom_abline(slope = coef(fit)["d.comp"],
              intercept = coef(fit)["(Intercept)"]) +
  scale_y_continuous("Competence scores for Democrats",
                     breaks = seq(-1, 1, by = 0.5), limits = c(-1, 1)) +
  scale_x_continuous("Democratic margin in vote shares",
                     breaks = seq(0, 1, by = 0.2), limits = c(0, 1)) 
```

## visualizing regression 2: `geom_point()` + `geom_smooth()`
```{r}
ggplot(data = face, mapping = aes(x = d.comp, y = diff.share)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  scale_y_continuous("Competence scores for Democrats",
                     breaks = seq(-1, 1, by = 0.5), limits = c(-1, 1)) +
  scale_x_continuous("Democratic margin in vote shares",
                     breaks = seq(0, 1, by = 0.2), limits = c(0, 1)) 
```

## visualizing regression 3: `geom_point()` + `geom_line()`
```{r include=FALSE}
fit.age2 <- lm(primary2006 ~ age + 
                 I(age^2) + messages +
                 age:messages + I(age^2):messages, 
               data = social.neighbor)
y.hat <- data_grid(social.neighbor, age, messages) %>%
  add_predictions(fit.age2)
```
```{r}
ggplot(y.hat, aes(x = age, y = pred)) +
  geom_line(aes(linetype = messages,
                color = messages)) +
  labs(color = "",
       linetype = "", y = "Predicted \nturnout rate",
       x = "Age") +
  xlim(20, 90) +
  theme(legend.position = "bottom")
```
