---
title: "Prediction 1 - tidyverse"
author: "Anna Yorozuya"
date: "June 14, 2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(stringr)
library(lubridate)
data("pres08", package = "qss")
data("pres12", package = "qss")
```


## `stringr` package - `str_c` function

- combining multiple objects into a character string
- each object will be separated by commas

```{r}
a <- 2
b <- 6
print(str_c(a, " times 3 is equal to ", b))
```

```{r}
data <- 1:4
for (i in seq_along(data)) {
    print(str_c("iteration", i))
  }
```


## `lubridate` package - `ymd` function
- converting strings in the format of `year-month-date` into a `date` object
- when subtracting a `date` object from another, the object will be a `difftime` object.
- the package has other commands, such as `mdy()` and `dmy()`

```{r}
# Converting an object into a date object
x <- ymd("2008-11-04")
y <- ymd("2008/9/1")
class(x)

# When subtracting a date object from another, 
# it will be a `difftime` object
(subtraction <- x - y)
class(subtraction)
```


## Combining objects
### tidyverse function `bind` (be careful!)
- `bind_cols(x, y)`: combine x and y by columns
- `bind_rows(x, y)`: combine x and y by rows

```{r}
# Combining two dataframes by columns
pres_bind_cols <- bind_cols(pres08, pres12)

# It looks good...?
colnames(pres08)
colnames(pres12)
colnames(pres_bind_cols)

# DC and Delware are in reverse order
pres_bind_cols %>%
  filter(state...2 != state...6)
```


### tidyverse function `mutate-joins` - better option!
- `inner_join(x, y, by = ): includes all rows in x and y
- `left_join (x, y, by = ): includes all rows in x
- `right_join (x, y, by = )`: includes all rows in y
- `full_join (x, y, by = )`: includes all rows in x or y

```{r}
## change the variable name for illustration
pres12 <- rename(pres12, state.abbrev = state)

## merging data sets using the variables of different names 
## and specifying the suffix
pres <- full_join(pres08, pres12,
                  by = c("state" = "state.abbrev"),
                  suffix = c("_08", "_12"))
colnames(pres)
```

