---
title: "Prediction 1 - tidyverse"
author: "Anna Yorozuya"
institute: "University of Tokyo"
date: "June 14, 2022"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
---

## Table of Contents
- `stringr` package
- `lubridate` package
- Combining objects
- Today's in-class assignment: `betting-markets`


## `stringr` package
### `stringr` package
- a package in tidyverse for manipulating strings
- here we use the function `str_c()`
- for more information, see [here](https://r4ds.had.co.nz/strings.html)

### `str_c` function
- combining multiple objects into a character string
- each object will be separated by commas

### Example
```{r include = FALSE}
library(tidyverse)
```
```{r}
a <- 2
b <- 6
print(str_c(a, " times 3 is equal to ", b))
```

## `lubridate` package
### `lubridate` package
- a package in tidyverse for dates-times object
- for more information, see [here](https://r4ds.had.co.nz/dates-and-times.html)

### `ymd` function
- converting strings in the format of `year-month-date` into a `date` object
- when subtracting a `date` object from another, the object will be a `difftime` object.

```{r include = FALSE}
library(lubridate)
```

## `lubridate` package - Example
```{r}
# Converting an object into a date object
x <- ymd("2008-11-04")
y <- ymd("2008/9/1")
class(x)

# When subtracting a date object from another, 
# it will be a `difftime` object
subtraction <- x - y
class(subtraction)
```


## Combining objects (data frame, vector, or matrix)
### base R function
- `cbind(x, y)`: combine x and y by columns
- `rbind(x, y)`: combine x and y by rows
- `merge(x, y, by = )`: merging two data frames by common columns or row names

### tidyverse function `bind` (be careful!)
- `bind_cols(x, y)`: combine x and y by columns
- `bind_rows(x, y)`: combine x and y by rows

### tidyverse function `mutate-joins`
- `inner_join(x, y, by = )`: includes all rows in x and y
- `left_join (x, y, by = )`: includes all rows in x
- `right_join (x, y, by = )`: includes all rows in y
- `full_join (x, y, by = )`: includes all rows in x or y

## References
- "Quantitative Social Science: An Introduction" - Kosuke Imai
- "Quantitative Social Science: An Introduction in Tidyverse" - Kosuke Imai and Nora Webb Williams
- [R for data science](https://r4ds.had.co.nz) - H.Wickham and G.Grolemund
- [stringr package](https://stringr.tidyverse.org)
- [lubridate package](https://lubridate.tidyverse.org)
