---
title: "Prediction 2 - tidyverse"
author: "Anna Yorozuya"
date: "June 16, 2022"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(modelr)
```

## Let's visualize the `lm()` outputs!

```{r}
data("face", package = "qss")
face <- mutate(face,
                d.share = d.votes / (d.votes + r.votes),
                r.share = r.votes / (d.votes + r.votes),
                diff.share = d.share - r.share)
face_fit <- lm(diff.share ~ d.comp, data = face)
face_fit
```

```{r}
data("florida", package = "qss")
florida_fit <- lm(Buchanan00 ~ Perot96, data = florida)
florida_fit
```

```{r}
data("social", package = "qss")
social_fit <- lm(primary2006 ~ messages, data = social)
social_fit
```


## `broom` package

`tidy()`: returns a data frame in which each row is a coefficient
```{r}
tidy(face_fit)
```

`glance()`: returns a one-row data frame summary of the model
```{r}
glance(face_fit)
```

- `augment()`: returns the original data with fitted values, residuals, and other observation level stats from the model appended to it.
```{r}
augment(face_fit) 
```

## `modelr` package

- `add_predictions()`: add the predictions to the original data
- `add_residuals()`: add the residuals to the original data
```{r}
florida_model <- florida %>%
  add_predictions(florida_fit) %>%
  add_residuals(florida_fit)
head(florida_model)
```

- `data_grid()`: create a data set containing every unique combination of the specified columns from the old data set.
```{r}
unique_messages <- social %>%
  data_grid(messages) %>%
  add_predictions(social_fit)
unique_messages
```

