---
title: "Measurement 1 tidyverse codes"
author: "Anna Yorozuya"
date: "6/2/2022"
output: html_document
---

```{r include = FALSE}
library(tidyverse)
data(afghan, package = "qss")
```

# Handling (missing) data

## Reordering rows with `arrange()`
reordering the rows from low to high (low to high with `desc()`)

```{r}
# create a data set
afghan_age <- afghan %>%
  select(province, district, age)

# reorder rows
arrange(afghan_age, age)
arrange(afghan_age, desc(age))
```


## Remove observations with NAs with `drop_na()`
Remove all observations with at least one missing value from a data frame.
**Be sure to check in advance how many NAs the data frame contains!**

```{r}
# check how many NAs will be omitted!

# original
nrow(afghan) 

# NAs omitted (without specifying columns)
afghan %>%
  drop_na() %>%
  nrow()

# NAs in income omitted
afghan %>%
  drop_na(income) %>%
  nrow() 
```


# Data visualization

```{r include = FALSE}
data("congress", package = "qss")

# create a dataset of the 80th congress
mydata <- filter(congress, congress == 80)
```


## Basic syntax (scatterplot)
````{r}
ggplot(data = abc) + 
  geom_point(aes(x = xxx, y = yyy))
```


## A simple scatterplot
````{r}
# plot the DW-nominate scores
ggplot(data = mydata) + 
  geom_point(aes(x = dwnom1,  ## dwnom1 = economic liberalism/conservatism
                 y = dwnom2)) ## dwnom2 = racial liberalism/conservatism

```

Compare it with the base R plot!
```{r}
plot(mydata$dwnom1, mydata$dwnom2,
     pch = 20)
```

Below, I will add different components to this scatterplot!


## how to combinate multiple plots
```{r}
ggplot(data = mydata) + 
  geom_point(aes(x = dwnom1, y = dwnom2)) +
  geom_vline(xintercept = 0, # add a vertical line
             color = "pink", linetype = "dashed") + # specify color and linetype
  geom_hline(yintercept = 0, # add a horizontal line
             color = "orange", linetype = "dashed") # specify color and linetype
```


## how to specify colors and shapes
```{r}
ggplot(data = mydata) + 
  geom_point(aes(x = dwnom1, y = dwnom2,
                 shape = party, color = party)) +  
  scale_color_manual(values = c(Democrat = "blue",
                                 Republican = "red",
                                 Other = "green")) +
  scale_shape_manual(values = c(Democrat = "square",
                                 Republican = "triangle",
                                 Other = "circle")) 
```


## how to add labels
```{r}
ggplot(data = mydata) + 
  geom_point(aes(x = dwnom1, y = dwnom2)) +
  scale_y_continuous("Racial liberalism/conservatism", limits = c(-1.5, 1.5)) +
  scale_x_continuous("Economic\n liberalism/conservatism", limits = c(-1.5, 1.5)) +
  coord_fixed()
```

## Let's put everything together!
```{r}
ggplot(data = mydata) +
  geom_point(aes(x = dwnom1, y = dwnom2,
                 shape = party, color = party)) +
  scale_color_manual(values = c(Democrat = "blue", # specify the color
                                 Republican = "red",
                                 Other = "green")) +
  scale_shape_manual(values = c(Democrat = "square", # specify the shape
                                 Republican = "triangle",
                                 Other = "circle")) +
  geom_vline(xintercept = 0, # add a vertical line
              color = "pink", linetype = "dashed") +
  geom_hline(yintercept = 0, # add a horizontal line
              color = "orange", linetype = "dashed") +
  scale_y_continuous("Racial liberalism/conservatism", limits = c(-1.5, 1.5)) +
  scale_x_continuous("Economic\n liberalism/conservatism", limits = c(-1.5, 1.5)) +
  coord_fixed()
```



