---
title: "Measurement 1 - tidyverse"
author: "Anna Yorozuya"
institute: "University of Tokyo"
date: "June 2, 2022"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
---

```{r include = FALSE}
library(tidyverse)
data(afghan, package = "qss")
```

## Table of Contents

- Handling (missing) data
- Introduction to `ggplot2`
- Today's in-class assignment: `gay-marriage-revisited`



# Handling (missing) data
## Handling (missing) data

### `arrange()`: reordering data
reordering the rows from low to high (low to high with `desc()`)
\tiny
```{r}
afghan_age <- select(afghan, province, district, age)
head(arrange(afghan_age, age))
head(arrange(afghan_age, desc(age)))
```

## Handling (missing) data

### `drop_na()`: listwise deletion (base R: `na_omit()`)
Remove all observations with at least one missing value from a data frame.\
```{r}
num1 <- c(2, 4, 3, NA)
num2 <- c(5, NA, 9, 8)
num <- tibble(num1, num2)
drop_na(num)
```


## Example for `arrange()`

\small

```{r}
## Table for non-missing values of ISAF and Taliban
afghan %>%
  filter(!is.na(violent.exp.ISAF), !is.na(violent.exp.taliban)) %>% 
  group_by(violent.exp.ISAF, violent.exp.taliban) %>%
  count() %>%
  ungroup() %>%
  mutate(prop = n / sum(n)) %>%
  arrange(prop) # compare to arrange(desc(prop))
```

\normalsize

## Example for `drop_na()`

\small

```{r}
# check how many NAs will be omitted!
nrow(afghan) # original
afghan.sub.2 <- drop_na(afghan)
nrow(afghan.sub.2) # NAs omitted
afghan %>%
  drop_na(income) %>%
  nrow() # NAs in income omitted
```

\normalsize

# Introduction to `ggplot2`
## Introduction to `ggplot2`
### What is `ggplot2`?
- A package in tidyverse, which allows visualization of data in a more intuitive way.\
- To make a plot, you assign the data and aesthetics (mapping) first, and add layers to tell the `ggplot` function what you want the figure to look like.
- The package name is **ggplot2**, while the function is **`ggplot()`**.\

### The advantage of using `ggplot2`
- Intuitive: very simple grammar
- Flexibility: you can build everything with the grammar
- Very nice-looking graphs!


## Basics of `ggplot2`
### Basic syntax
| ggplot(data = **DATA**) + 
|   **GEOM_FUNCTION**(mapping = aes(**MAPPINGS**)) + 
|   **ADDITIONAL_FUNCTIONS**
|       * Add the components with "`+`"

### Elements
- **DATA**: specify the dataset to use in the graph
- **GEOM_FUNCTION**: starting from "`geom_`", where you specify the types of figures such as bar plot or histogram.
- **MAPPINGS**: defines how variables in your dataset are mapped to visual properties. Commonly used arguments are `x` and `y` to specify which variables to map to each axis.
- **ADDITIONAL_FUNCTIONS**: additional layers


## `ggplot2`: `GEOM_FUNCTION`
### Plots
- `geom_point()`: scatterplot
- `geom_histogram()`: histogram
- `geom_bar()`: bar plot
- `geom_boxplot()`: box plot
- `geom_line()`: line chart
- `geom_smooth()`: smooth line, mainly for regression
- `geom_ribbon()`: show confidence intervals

### Line
- `geom_abline()`: intercept, slope
- `geom_hline()`: yintercept
- `geom_vline()`: xintercept


## `ggplot2`: `aes(MAPPINGS)`
### Aesthetics
- `x = variable`: values for x axis
- `y = variable`: values for y axis
- `color = variable`: assign unique color for each value of the variable
- `fill = variable`: assign the unique color to fill in for each value
- `size = variable`: assign unique size for each value of the variable
- `alpha = variable`: control the level of transparency for each value
- `shape = variable`: change the shape of points in `geom_point`
- `position = `: `fill` for stacking, `dodge` for avoiding overlapping, `jitter` for solving overplotting\

### *All of those above should be within `aes()`
- If it goes outside of `aes()`, the above arguments will be applied to all the variables, regardless of the value, unless specified.


## `ggplot2`: `ADDITIONAL_FUNCTIONS`
### Scales
| Map the data values to visual values
|    **`scale_!_!!`**, where
|         **!**: aesthetic to adjust (x, y, fill, etc.)
|         **!!**: prepackaged scale to use 

- `scale_x_discrete()`: set discrete values for visualization
- `scale_x_continuous()`: set continuous values for visualization
- `scale_fill_discrete()`: fill the plot with discrete values

### Labels
- `labs(title = "", x = "", y = "")`: label for x-axis, y-axis, and title
- `ylab("label")`: label for y-axis
- `xlab("label")`: label for x-axis
- `ggtitle("title"): title

## `ggplot2`: `ADDITIONAL_FUNCTIONS`
### Limits
- `xlim()`: limits for x-axis
- `ylim()`: limits for y-axis

### Themes
- `theme_classic()`
- `theme_bw()`: white background with grid lines
- `theme_gray()`: grey background (default)
- `theme_void()`: empty theme

## `ggplot2`: `ADDITIONAL_FUNCTIONS`
### Facets
- `facet_wrap( ~ a)`: facet by a single variable (a in the example)
- `facet_grid(b ~ c)`: facet by two variables 

### Coordinate systems
- `coord_flip()`: switch x and y axis
- `coord_quickmap()`: set the aspect ratio correctly for maps
- `coord_fixed()`: fix the aspect ratio to square

## Example: Scatterplot

\tiny

```{r include = FALSE}
library(gridExtra)
data("congress", package = "qss")
```
```{r fig.show="hold", out.width="30%", fig.align='center'}
ggplot(data = filter(congress, congress == 80),
       aes(x = dwnom1, y = dwnom2)) +
  geom_point(aes(shape = party, color = party),
             show.legend = FALSE) +
  scale_color_manual(values = c(Democrat = "blue",
                                 Republican = "red",
                                 Other = "green")) +
  scale_shape_manual(values = c(Democrat = "square",
                                 Republican = "triangle",
                                 Other = "circle")) +
  scale_y_continuous("Racial liberalism/conservatism", limits = c(-1.5, 1.5)) +
  scale_x_continuous("Economic\n liberalism/conservatism", limits = c(-1.5, 1.5)) +
  coord_fixed()
```
- `scale_color_manual()`: specify which colors are used for which value
- `scale_shape_manual()`: specify which shape is used for which value
- `scale_y_continuous()`/ `scale_x_continuous()`: add title, change the limits
- `coord_fixed()`: squared aspect ratio

## Example: Histogram (basic)
\tiny
```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(x = age)) + # the data and initial aes() 
  geom_histogram(aes(y = ..density..), # histogram, additional aes()
                 binwidth = 5, # how wide for each bin
                 boundary = 0) + # bin position 
  scale_x_continuous(breaks = seq(20, 80, by = 10)) +
  labs(title = "Distribution of respondent's age",
       y = "Density", x = "Age") +
  theme_classic()
```
- `aes(y = ..density..)`: y-axis shows the density, not the count
- `aes(binwidth = 5)`: set the width of each bin
- `aes(boundary = 0)`: the position of bins
- `scale_x_continuous()`: change the ticks of x-axis

## Example: Histogram (advanced)
\tiny
```{r  fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(x = educ.years, y = ..density..)) +
  geom_histogram(binwidth = 1, center = 0) +
  geom_vline(xintercept = median(afghan$educ.years)) +
  annotate(geom = "text", x = median(afghan$educ.years),
           y = 0.4,
           label = "median",
           hjust = -0.1) +
  labs(title = "Distribution of respondent's education",
       x = "Years of education",
       y = "Density")
```
- `geom_vline()`: add a vertical line
- `annotate()`: add text to the plot. specify the position and text


## How to save/print graphs
### `ggsave`
- `ggsave(path, filename, extension)`\
- for example, if you want to wave the figure as a pdf in the result_fingures directory,\
`ggsave("results_figures/education_by_province.pdf")`

### `gridExtra`
- save multiple plots into a single file
- first, load the package with `library(gridExtra)`
- use the `grid_arrange()`


## Example: gridExtra
\tiny
```{r fig.show="hold", out.width="50%", fig.align='center'}
library(gridExtra)
## The age histogram
age_hist <- ggplot(afghan, aes(x = age)) +
  geom_histogram(aes(y = ..density..), binwidth = 5, boundary = 0) +
  scale_x_continuous(breaks = seq(20, 80, by = 10)) +
  labs(title = "Distribution of \nrespondent's age", y = "Age", x = "Density")
## The education histogram
educ_hist <- ggplot(afghan, aes(x = educ.years, y = ..density..)) +
  geom_histogram(binwidth = 1, center = 0) +
  geom_vline(xintercept = median(afghan$educ.years)) +
  annotate(geom = "text", x = median(afghan$educ.years), y = 0.4, label = "median", hjust = -0.1) +
  labs(title = "Distribution of \nrespondent's education", x = "Years of education", y = "Density")
## Put the plots side-by-side
grid.arrange(age_hist, educ_hist, ncol = 2)
```


## References
- "Quantitative Social Science: An Introduction" - Kosuke Imai
- "Quantitative Social Science: An Introduction in Tidyverse" - Kosuke Imai and Nora Webb Williams
- [R for data science](https://r4ds.had.co.nz) - H.Wickham and G.Grolemund


