---
title: "Measurement 2 - tidyverse"
author: "Anna Yorozuya"
institute: "University of Tokyo"
date: "June 7, 2022"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
---

```{r include = FALSE}
library(tidyverse)
data(afghan, package = "qss")
```


## Table of Contents
- How to save/print graphs
- Review of `ggplot2`
- `tidymodels` package
- Today's in-class assignment: `political-efficacy`


## How to save/print graphs
### `ggsave`
- `ggsave(path, filename, extension)`\
- for example, if you want to save the figure as a pdf in the result_figures directory,\
`ggsave("results_figures/education_by_province.pdf")`

### `gridExtra`
- save multiple plots into a single file
- first, load the package with `library(gridExtra)`
- use the `grid_arrange()`


# Review: `ggplot2`
## Example: Bar plot (basic)

\tiny

```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(data = afghan, # Tell R what data to use
    aes(x = as.factor(violent.exp.ISAF))) + # specify the x-axis 
  geom_bar(aes(y = stat(prop), # add a bar plot layer
           group = 1)) + 
  scale_x_discrete(labels = c("No Harm","Harm","Nonresponse")) + 
  ylab("Proportion of Respondents") + # Add a label to y-axis 
  xlab("Response Category") + # Add a label to the x-axis 
  ggtitle("Civilian Victimization by the ISAF") # Add a title
```
- `aes(y = stat(prop))`: the y-axis shows the proportion, not the count. this is equivalent to `aes(y = ..prop..)`
- `aes(group = 1)`: plot the proportion of the total
- `scale_x_discrete()`: indicate which value to display in the x-axis

## Example: Bar plot (advanced)
\tiny
```{r include = FALSE}
afghan_reshape <- afghan %>%
  pivot_longer(violent.exp.ISAF:violent.exp.taliban,
               names_to = "harming_group",
               values_to = "harm")
```

```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(data = afghan_reshape,  
       aes(x = as.factor(harm))) +
geom_bar(aes(y = ..prop.., fill = harming_group,
           group = harming_group),
           position = "dodge") +
  scale_x_discrete(labels = c("No Harm","Harm","Nonresponse")) +
  scale_fill_discrete(name = "Harming Group", labels = c("ISAF", "Taliban")) +
  ylab("Proportion of Respondents") +
  xlab("Response Category") +
  ggtitle("Civilian Victimization")
```
- `position = "dodge"`: avoid overlapping, places bars side by side
- `scale_fill_discrete()`: change the labels on the legend for the bar colors


## Example: Boxplot (basic)
```{r  fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(y = age)) +
  geom_boxplot() +
  labs(y = "Age", x = "", title = "Distribution of Age")
```

## Example: Boxplot (advanced)
```{r  fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(y = educ.years, x = province)) +
  geom_boxplot() +
  labs(y = "Year of Education", x = "Province", title = "Distribution of Age")
```
- `aes(x = province)`: create boxplot for each value of province


## `tidymodels` package
### What is `tidymodels`?
- "a collection of packages for modeling and machine learning using tidyverse principles" (official description of this package)
- the `tidy()` function here is contained in the `broom` package as well

### function `tidy()`: convert the output of a model into a tibble
- takes the output of an R model (such as `kmeans()`) as an argument

## References
- "Quantitative Social Science: An Introduction" - Kosuke Imai
- "Quantitative Social Science: An Introduction in Tidyverse" - Kosuke Imai and Nora Webb Williams
- [R for data science](https://r4ds.had.co.nz) - H.Wickham and G.Grolemund
- [Tidymodels package](https://www.tidymodels.org) 

