---
title: "Measurement tidyverse"
author: "Anna Yorozuya"
institute: "University of Tokyo"
date: "June 2022"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "beaver"
---

```{r include = FALSE}
library(tidyverse)
data(afghan, package = "qss")
```
# Measurement 1 (June 2)
## Table of Contents
- Handling missing data
- Introduction to `ggplot2`
- Today's in-class assignment: `gay-marriage-revisited`


# Handling missing data
## Handling missing data
### `arrange()` for reordering data
- Description: reordering the rows from low to high
- Arguments: add `desc()` for reordering from high to low

### `drop_na()` for listwise deletion
- Description: remove all observations(rows) with at least one missing value from a data frame
- Arguments: the name of data frame
- Corresponding base R function: `na_omit()`
- **warning: check how many NAs are in the data frame before applying listwise deletion!**


## Example for `arrange()`

\small

```{r}
## Table for non-missing values of ISAF and Taliban
afghan %>%
  filter(!is.na(violent.exp.ISAF), !is.na(violent.exp.taliban)) %>% 
  group_by(violent.exp.ISAF, violent.exp.taliban) %>%
  count() %>%
  ungroup() %>%
  mutate(prop = n / sum(n)) %>%
  arrange(prop) # compare to arrange(desc(prop))
```

\normalsize

## Example for `drop_na()`

\small
```{r}
nrow(afghan) # original
afghan.sub.2 <- drop_na(afghan)
nrow(afghan.sub.2) # NAs omitted
afghan %>%
  drop_na(income) %>%
  nrow() # NAs in income omitted
```
\normalsize

# Introduction to `ggplot2`
## Introduction to `ggplot2`
### What is `ggplot2`?
- A package in tidyverse, which allows visualization of data in a more intuitive way.\
- To make a plot, you assign the data and aesthetics (mapping) first, and add layers to tell the `ggplot` function what you want the figure to look like.
- The package name is **ggplot2**, while the function is **`ggplot()`**.\

### The advantage of using `ggplot2`
- Intuitive: very simple grammar
- Flexibility: you can build everything with the grammar
- Very nice-looking graphs!


## Basics of `ggplot2`
### Basic syntax
| ggplot(data = DATA) + 
|   GEOM_FUNCTION(mapping = aes(MAPPINGS)) + 
|   ADDITIONAL_FUNCTIONS
**Add the components with `+`**

### Elements
- DATA: specify the dataset to use in the graph
- GEOM_FUNCTION: starting from`geom_`. specify the types of figures such as bar plot or histogram.
- MAPPINGS: defines how variables in your dataset are mapped to visual properties. commonly used arguments are `x` and `y` to specify which variables to map to each axis.
- ADDITIONAL_FUNCTIONS


## `ggplot2`: `GEOM_FUNCTION`
### Plots
- `geom_point()`: scatterplot
- `geom_histogram()`: histogram
- `geom_bar()`: bar plot
- `geom_boxplot()`: box plot
- `geom_line()`: line chart
- `geom_smooth()`: smooth line, mainly for regression

### Line
- `geom_abline()`: intercept, slope
- `geom_hline()`: yintercept
- `geom_vline()`: xintercept


## `ggplot2`: `aes(MAPPINGS)`
### Aesthetics
- `x = variable`: value for x axis
- `y = variable`: value for y axis
- `color = variable`: assign unique color for each value of the variable
- `fill = variable`: assign the unique color to fill in for each value
- `size = variable`: assign unique size for each value of the variable
- `alpha = variable`: control the level of transparency for each value
- `shape = variable`: change the shape of points in `geom_point`
- `position = `: `fill` for stacking, `dodge` for avoiding overlapping, `jitter` for solving overplotting
**All of those above should be within `aes()`**
If it goes outside of `aes()`, the above arguments will be applied to all the variables, regardless of the value, unless specified.


## `ggplot2`: `ADDITIONAL_FUNCTIONS`
### Scales
Map the data values to visual values
| `scale_*_**`, where
|    *: aesthetic to adjust (x, y, fill, etc.)
|    **: prepackaged scale to use 
- `scale_x_discrete()`: set the discrete values for visualization
- `scale_x_continuous()`: set the continuous values for visualization
- `scale_fill_discrete()`: fill the plot with discrete values

### Labels
- `labs(title = "", x = "", y = "")`: label for x-axis, y-axis, and title
- `ylab("label")`: label for y-axis
- `xlab("label")`: label for x-axis
- `ggtitle("title"): title

## `ggplot2`: `ADDITIONAL_FUNCTIONS`
### Limits
- `xlim()`: limits for x-axis
- `ylim()`: limits for y-axis

### Themes
- `theme_classic()`
- `theme_bw()`: white background with grid lines
- `theme_gray()`: grey background (default)
- `theme_void()`: empty theme

## `ggplot2`: `ADDITIONAL_FUNCTIONS`
### Facets
- `facet_wrap( ~ a)`: facet by a single variable (a in the example)
- `facet_grid(b ~ c)`: facet by two variables 

### Coordinate systems
- `coord_flip()`: switch x and y axis
- `coord_quickmap()`: set the aspect ratio correctly for maps
- `coord_fixed()`: fix the aspect ratio to square

## Example: Scatterplot

\tiny 

```{r include = FALSE}
library(gridExtra)
data("congress", package = "qss")
```
```{r fig.show="hold", out.width="30%", fig.align='center'}
ggplot(data = filter(congress, congress == 80), aes(x = dwnom1, y = dwnom2)) +
  geom_point(aes(shape = party, color = party), show.legend = FALSE) +
  scale_color_manual(values = c(Democrat = "blue",
                                 Republican = "red",
                                 Other = "green")) +
  scale_shape_manual(values = c(Democrat = "square",
                                 Republican = "triangle",
                                 Other = "circle")) +
  scale_y_continuous("Racial liberalism/conservatism", limits = c(-1.5, 1.5)) +
  scale_x_continuous("Economic\n liberalism/conservatism",limits = c(-1.5, 1.5)) +
  ggtitle("80th Congress") +
  coord_fixed()
```

- `scale_color_manual()`: specify which colors are used for which value
- `scale_shape_manual()`: specify which shape is used for which value
- `scale_y_continuous()`/ `scale_x_continuous()`: add title, change the limits
- `coord_fixed()`: squared aspect ratio

## Example: Histogram (basic)

\tiny

```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(x = age)) + # the data and initial aes() 
  geom_histogram(aes(y = ..density..), # histogram, additional aes()
                 binwidth = 5, # how wide for each bin
                 boundary = 0) + # bin position 
  scale_x_continuous(breaks = seq(20, 80, by = 10)) +
  labs(title = "Distribution of respondent's age",
       y = "Density", x = "Age") +
  theme_classic()
```
- `aes(y = ..density..)`: y-axis shows the density, not the count
- `aes(binwidth = 5)`: set the width of each bin
- `aes(boundary = 0)`: the position of bins
- `scale_x_continuous()`: change the ticks of x-axis

## Example: Histogram (advanced)

\tiny 

```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(x = educ.years, y = ..density..)) +
  geom_histogram(binwidth = 1, center = 0) +
  geom_vline(xintercept = median(afghan$educ.years)) +
  annotate(geom = "text", x = median(afghan$educ.years),
           y = 0.4,
           label = "median",
           hjust = -0.1) +
  labs(title = "Distribution of respondent's education",
       x = "Years of education",
       y = "Density")
```
- `geom_vline()`: add a vertical line
- `annotate()`: add text to the plot. specify the position and text


## How to save/print graphs
### `ggsave`
- `ggsave(path, filename, extension)`\
- for example, if you want to wave the figure as a pdf in the result_fingures directory,\
`ggsave("results_figures/education_by_province.pdf")`

### `gridExtra`
- save multiple plots into a single file
- first, load the package with `library(gridExtra)`
- use the `grid_arrange()`

## Example: gridExtra

\tiny

```{r}
library(gridExtra)
## The age histogram
age_hist <- ggplot(afghan, aes(x = age)) +
  geom_histogram(aes(y = ..density..), binwidth = 5, boundary = 0) +
  scale_x_continuous(breaks = seq(20, 80, by = 10)) +
  labs(title = "Distribution of \nrespondent's age", y = "Age", x = "Density")
```
```{r fig.show="hold", out.width="50%", fig.align='center'}
## The education histogram
educ_hist <- ggplot(afghan, aes(x = educ.years, y = ..density..)) +
  geom_histogram(binwidth = 1, center = 0) +
  geom_vline(xintercept = median(afghan$educ.years)) +
  annotate(geom = "text", x = median(afghan$educ.years),y = 0.4, label = "median", hjust = -0.1) +
  labs(title = "Distribution of \nrespondent's education", x = "Years of education", y = "Density")
```
```{r fig.show="hold", out.width="50%", fig.align='center'}
## Put the plots side-by-side
grid.arrange(age_hist, educ_hist, ncol = 2)
```


## memo (delete before publishing)
filter, summarize(corr, use), select
ggplot, geom_point, geom_histogram(bins), xlab, ggtitle
mutate(if_else, is.na, as.double)
qqplot


# Measurement 2 (June 7)
## Table of Contents
- Review of `ggplot2`

- `tidymodels` package


- Today's in-class assignment: `political-efficacy`


## Review: `ggplot2`


## Example: Bar plot (basic)

\tiny

```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(data = afghan, # Tell R what data to use
    aes(x = as.factor(violent.exp.ISAF))) + # specify the x-axis 
  geom_bar(aes(y = ..prop.., # add a bar plot layer
           group = 1)) + 
  scale_x_discrete(labels = c('No Harm','Harm','Nonresponse')) + 
  ylab("Proportion of Respondents") + # Add a label to y-axis 
  xlab("Response Category") + # Add a label to the x-axis 
  ggtitle("Civilian Victimization by the ISAF") # Add a title
```
- `aes(y = ..prop..)`: the y-axis shows the proportion, not the count
- `aes(group = 1)`: plot the proportion of the total
- `scale_x_discrete()`: indicate which value to display in the x-axis

## Example: Bar plot (advanced)

\tiny

```{r include = FALSE}
afghan_reshape <- afghan %>%
  pivot_longer(violent.exp.ISAF:violent.exp.taliban,
               names_to = "harming_group",
               values_to = "harm")
```

```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(data = afghan_reshape,  
       aes(x = as.factor(harm))) +
geom_bar(aes(y = ..prop.., 
           fill = harming_group,
           group = harming_group),
           position = "dodge") +
  scale_x_discrete(labels = c('No Harm','Harm','Nonresponse')) +
  scale_fill_discrete(name = "Harming Group", labels = c("ISAF", "Taliban")) +
  ylab("Proportion of Respondents") +
  xlab("Response Category") +
  ggtitle("Civilian Victimization")
```
- `position = "dodge"`: avoid overlapping, places bars side by side
- `scale_fill_discrete()`: change the labels on the legend for the bar colors


## Example: Boxplot (basic)
```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(y = age)) +
  geom_boxplot() +
  labs(y = "Age", x = "", title = "Distribution of Age")
```

## Example: Boxplot (advanced)
```{r fig.show="hold", out.width="50%", fig.align='center'}
ggplot(afghan, aes(y = educ.years, x = province)) +
  geom_boxplot() +
  labs(y = "Year of Education", x = "Province", title = "Distribution of Age")
```
- `aes(x = province)`: create boxplot for each value of province


## `tidymodels` package
### What is `tidymodels`?
- a collection of packages for modeling and machine learning using tidyverse principles
- the `tidy()` function here is contained in the`broom` package as well

### function `tidy()`
- description: convert the output of a model into a tibble
- argument: the output of an R model (such as `kmeans()`)


## memo (delete before publishing)
group_by, ungroup, summarize, mutate (~), filter, 
ggplot, geom_bar (position, stat), labs, geom_histogram, geom_vline, annotate, labs
qqplot

# Reference
## Reference

- [R for data science](http://r4ds.had.co.nz)

- [R cheatsheet - visualization](https://www.rstudio.com/resources/cheatsheets/)

