---
title: 'Chapter 2: Causality'
author: "Claire Liow"
date: "5/31/2022"
output:
  beamer_presentation:
    slide_level: 2
    toc: yes
    theme: Madrid
    colortheme: beaver
  pdf_document:
    toc: yes
subtitle: Data Transformation with Tidyverse Functions
institute: University of Tokyo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## load packages
library(tidyverse)
library(qss)
## load data
resume <- read_csv("~/Dropbox/QSS/qss-inst-tidyverse/Causality/causality_tidy_files/data/resume.csv")
```

# Aggregate data with **`group_by()`**
## What does it do?
- Group data for downstream analysis
- Commonly used with `summarize` and `mutate`

```{r, eval = FALSE, echo = TRUE}
## count the observation in selected variables
race.call.summary <- resume %>%
  group_by(race, call) %>% 
  count() 

race.call.summary

## calculate callback rates 
callback_by_race <- resume %>%
  group_by(race, sex) %>%
  summarize(callback_rate = mean(call)) 

callback_by_race
``` 

# Summarize data with **`summarize()`**
## What does it do?
- Group $\rightsquigarrow$ Summarize
- Collapse each group into a single row summary

```{r, eval = FALSE, echo = TRUE}
## calculate callback rate by race
resume %>%
group_by(race) %>%
summarize(callback = mean(call))

## calculate callback rate by race and sex
resume %>%
group_by(race, sex) %>%
summarize(callback = mean(call))
```

# Reshape the data with **`pivot_wider()`**
## What does it do?
- Increasing the number of columns
- Decreasing the number of rows

```{r, eval = FALSE, echo = TRUE}
## tidyverse
## without pivoting the data
resume %>%
group_by(race, sex) %>%
summarize(callback = mean(call))

## after pivoting the data
resume %>%
group_by(race, sex) %>%
summarize(callback = mean(call)) %>%
pivot_wider(names_from = race, values_from = callback)
```

# Summary
## Today's tidyverse functions:
- **`group_by()`** to aggregate data
- **`summarize()`** to summarize data
- **`pivot_wider()`** to reshape data

## Reference
- [Quantitative Social Science: An Introduction in tidyverse](https://press.princeton.edu/books/hardcover/9780691222271/quantitative-social-science)