---
title: "Chapter 2: Causality"
subtitle: "Data Transformation with Tidyverse"
author: "Sho Miyazaki"
institute: "Keio University"
date: "5/26/2022"
output: 
  beamer_presentation:
    slide_level: 2
    toc: true
    theme: "Madrid"
    colortheme: "beaver"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 100)
library(knitr)
library(cowplot)
```

# Introduction

## Let's get started with Data
\small
```{r warning=FALSE, message=FALSE}
## load packages
library(tidyverse)
library(qss)
## load data
resume <- read_csv("data/resume.csv")
# check data 
resume
```

## Today's Goal
### Combine functions to get informative output
\small
```{r warning=FALSE, message=FALSE}
racial_gaps_by_sex <- resume %>%
  group_by(race, sex) %>% 
    # using two variables to group the data 
  summarize(callback = mean(call)) %>% 
    # the callback rate for each group 
  pivot_wider(names_from = race, 
                # reshaping the data
              values_from = callback) %>%
  mutate(race_gap = white - black)
```
\large

```{r echo=FALSE}
kable(racial_gaps_by_sex)
```

## Tools
```{r fig.show="hold", out.width="40%", fig.align='center',echo=FALSE}
ggdraw() + 
  draw_image("figure/tidyr_logo.png") 
ggdraw() + 
  draw_image("figure/dplyr_logo.png") 
```

## `dplyr` from `Tidyverse`

```{r fig.show="hold", out.width="80%", fig.align='center', echo=FALSE}
ggdraw() + 
  draw_image("figure/dplyr.png") 
```
*Source: [RStudio](https://www.rstudio.com/resources/webinars/tidyverse-visualization-manipulation-basics/)*

# `dplyr` and `%>%`

## What is "`pipe %>% `" ?

```{r fig.show="hold", out.width="25%", fig.align='center', echo=FALSE}
ggdraw() + 
  draw_image("figure/pipe.png") 
```

- "a good way to pronounce %>% when reading code is “**then**”."

```{r fig.show="hold", out.width="48%", fig.align='center',echo=FALSE}
ggdraw() + 
  draw_image("figure/tidy_code.png") 
ggdraw() + 
  draw_image("figure/baser_code.png") 
```

\tiny
*Source: [R for Data Science](https://r4ds.had.co.nz/explore-intro.html)*

## Extract Rows (`filter`)

- `filter`: Return rows by name/number/etc.

```{r}
## subset data with black names
resume %>% 
  filter(race == "black")
```

## Extract Columns (`select`) 

- `select`: Return columns by name/number/etc.
```{r}
## Subset with sex and race columns
resume %>% 
  select(sex, race)
```

## Compute New Columns (`mutate`)

- `mutate`

\scriptsize
```{r}
## create a factor variable that takes one of the four values
resume %>%
  mutate(type = case_when(race == "black" & sex == "female" ~ "BlackFemale",
                          race == "black" & sex == "male" ~ "BlackMale",
                          race == "white" & sex == "female" ~ "WhiteFemale",
                          race == "white" & sex == "male" ~ "WhiteMale",
                          TRUE ~ "Other"))
```

## Compute Table Summaries (`summarise`)

```{r}
## callback rate for black female names
resume %>% 
  filter(race == "black" & sex =="female") %>%
  summarize(callback_rate = mean(call, na.rm = TRUE))
```

# Summary

## Overwhelmed? 
### Don't worry!
There are many resources you can use, and you don't have to memorize all the functions.

- QSS Textbook
  + Tidyverse Version is on `Perusall`
- Cheetsheets
  + Search "`tidyverse cheetsheets`"
  + https://www.rstudio.com/resources/cheatsheets/
- Online Resources 
  + Google "`tidyverse add column error`"
  + official reference page, stackoverflow, RPubs, etc. 

### Teaching Team
We are here for you!

## Let's practice!
```{r fig.show="hold", out.width="80%", fig.align='center', echo=FALSE}
ggdraw() + 
  draw_image("figure/dplyr.png") 
```
*Source: [RStudio](https://www.rstudio.com/resources/webinars/tidyverse-visualization-manipulation-basics/)*
