---
title: "Introduction to Tidyverse"
author: "Claire Liow"
institute: "University of Tokyo"
date: "5/24/2022"
output: 
  beamer_presentation:
    slide_level: 2
    toc: true
    theme: "Madrid"
    colortheme: "beaver"
---
# What is tidyverse?
```{r}
#![Alt text] (~/Desktop/tidyverse_packages.png)
```
- A collection of R packages 
- Support natural workflow of data analysis
- Data import, tidying, manipulation, visualization, programming

## Load tidyverse
```{r, eval=FALSE, echo=TRUE}
#install.packages(tidyverse)
library("tidyverse")
```

# Base R versus tidyverse syntax
## Base R versus tidyverse syntax
### Base R syntax
```{r, eval=FALSE, echo=TRUE}
UNpop <- read.csv("data/UNpop.csv")
class(UNpop)
head(UNpop)

UNpop$world.pop # access individual variable
UNpop[, "world.pop"] # extract the column "world.pop"
UNpop[1:3, "year"] # subset first three rows of year 
```

### tidyverse syntax
```{r, eval=FALSE, echo=TRUE}
UNpop.tidy <- read_csv("data/UNpop.csv")
class(UNpop.tidy)
glimpse(UNpop.tidy)

select(UNpop.tidy, world.pop) # extracting the world.pop variable
UNpop.tidy %>% # subset the first three rows of years
  select(year) %>%
  slice(1:3)
```

# Summary
## Summary
###Differences between Base R/tidyverse?
- Base R: Lots of $, [[]]
- Tidyverse: %>% (Forward pipe operator)

###When to use Base R/tidyverse?
- 

