---
title: "Chapter 1: Introduction"
subtitle: "What is tidyverse?"
author: "Claire Liow"
institute: "University of Tokyo"
date: "5/24/2022"
output: 
  beamer_presentation:
    slide_level: 2
    toc: true
    theme: "Madrid"
    colortheme: "beaver"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(width = 100)
library(tidyverse)
library(knitr)
library(cowplot)
```

# What is `R` and `tidyverse`?

## What is `R`?
A programing language!
```{r fig.show="hold", out.width="80%", fig.align='center'}
ggdraw() + 
  draw_image("~/Dropbox/QSS/qss-inst-tidyverse/Introduction/introduction_tidy_files/figure/R_as_a_language.png") 
```
*Source: [RStudio](https://www.rstudio.com/resources/webinars/tidyverse-visualization-manipulation-basics/)*

\newpage
## Then, what is `tidyverse`?
```{r fig.show = "hold", out.width = "30%", fig.align = 'center'}
ggdraw() + 
  draw_image("~/Dropbox/QSS/qss-inst-tidyverse/Introduction/introduction_tidy_files/figure/tidyverse_logo.png") 
```
- A collection of R packages 
- A dialect of the Base R programming language
- Support natural workflow of data analysis
- Data import, tidying, manipulation, visualization, programming

\newpage
## Then, what is `tidyverse`?
```{r fig.show = "hold", out.width = "80%", fig.align = 'center'}
ggdraw() + 
  draw_image("~/Dropbox/QSS/qss-inst-tidyverse/Introduction/introduction_tidy_files/figure/packages.png") 
```
*Source: [Slyvia Canelon](https://github.com/spcanelon/tour-of-the-tidyverse/blob/master/images/pptx/tidyverse.png)*

\newpage

# Differences between Base R and tidyverse
## Load tidyverse
```{r, eval = FALSE, echo = TRUE}
#install.packages(tidyverse)
library("tidyverse")
```

## Example
```{r echo = TRUE}
UNpop <- read.csv("~/Dropbox/QSS/qss-inst-tidyverse/Introduction/introduction_tidy_files/data/UNpop.csv")
UNpop
```

Let's calculate the \% of population increase from 1950!

\newpage

## Base R syntax
```{r echo = TRUE}
## calculate the ratio compared to 1950
UNpop$ratio <- UNpop$world.pop / UNpop$world.pop[1] 

## convert to percentage increase and round
UNpop$percent <- round((UNpop$ratio - 1) * 100, 1)
```

\newpage

## tidyverse syntax
```{r, echo = TRUE}
UNpop %>% 
  # calculate the ratio compared to 1950
  mutate(ratio = world.pop / first(world.pop), 
  # convert to percentage increase and round
  percent = round((ratio - 1) * 100, 1)) 
```

\newpage

## Same task, with different syntax
### `Base-R`
```{r eval = FALSE, echo = TRUE}
UNpop$ratio <- UNpop$world.pop / UNpop$world.pop[1] 
UNpop$percent <- round((UNpop$ratio - 1) * 100, 1)
```

### `tidyverse`
```{r eval = FALSE, echo = TRUE}
UNpop %>% 
  mutate(ratio = world.pop / first(world.pop),
         percent = round((ratio - 1) * 100, 1)) 
```

\newpage

# Summary
## Summary
### How to distinct between base R/tidyverse?
- Base R: Lots of **$** and **[[ ]]**
- Tidyverse: **%>%** (Forward pipe operator)

### Why tidyverse?
- Resemble human language 
- Intuitive and logical 
- Combination of both base R and tidyverse

```{r fig.show = "hold", out.width = "40%", fig.align = 'center'}
ggdraw() + 
  draw_image("~/Dropbox/QSS/qss-inst-tidyverse/Introduction/introduction_tidy_files/figure/pipe.png") 
```
*Source: [RStudio](https://www.rstudio.com/resources/webinars/tidyverse-visualization-manipulation-basics/)*

## Reference
- Quantitative Social Science: An Introduction to tidyverse
- [RStudio](https://www.rstudio.com/resources/webinars/tidyverse-visualization-manipulation-basics/)
- [Slyvia Canelon](https://github.com/spcanelon/tour-of-the-tidyverse/blob/master/images/pptx/tidyverse.png)
