---
title: "Sex Ratio and the Price of Agricultural Crops in China"
output: html_document
header-includes:
   - \usepackage{amsmath}
   - \usepackage{bm}
   - \newcommand{\E}{\mathbb{E}}
   - \newcommand{\V}{\mathbb{V}}
---

In this exercise, we consider the effect of a change in the price of
agricultural goods whose production and cultivation are dominated by
either men or women. 

This exercise is based on: Qian, Nancy. 2008. “[Missing Women and the Price of Tea in China: 
 The Effect of Sex-Specific Earnings on Sex 
 Imbalance.](http://dx.doi.org/10.1162/qjec.2008.123.3.1251)” *Quarterly 
 Journal of Economics* 123(3): 1251–85. 

Our data come from China, where centrally planned production targets during the Maoist
era led to changes in the prices of major staple crops. We focus here
on tea, the production and cultivation of which required a large
female labor force, as well as orchard fruits, for which the labor
force was overwhelmingly male.  We use price increases brought on by
government policy change in 1979 as a proxy for increases in
sex-specific income, and ask the following question: Do changes in
sex-specific income alter the incentives for Chinese families to have
children of one gender over another? The CSV data file,
`chinawomen.csv`, contains the variables shown in the table below, with each observation representing a particular Chinese county in a given year. Note that `post` is
an indicator variable that takes takes 1 in a year following the
policy change and 0 in a year before the policy change.
   
------------------------------------------------------------------------------
 Name                Description
 ------------------- ---------------------------------------------------------
 `birpop`            Birth population in a given year
 
 `biryr`             Year of cohort (birth year)
 
 `cashcrop`          Amount of cash crops planted in county
 
 `han`               Fraction ethnically Han
 
 `orch`              Amount of orchard-type crops planted in county
 
 `teasown`           Amount of tea sown in county 
 
 `sex`               Proportion of males in birth cohort
 
 `post`              Indicator variable for introduction of price reforms
------------------------------------------------------------------------------

## Question 1

We begin by examining sex ratios in the post-reform period (that
  is, the period after 1979) according to whether or not tea crops
  were sown in the region.  Estimate the mean sex ratio in 1985, which
  we define as the proportion of male births, separately for
  tea-producing and non-tea-producing regions.  Compute the 95\%
  confidence interval for each estimate by assuming independence
  across counties within a year (We will maintain this assumption
  throughout this exercise).  Furthermore, compute the
  difference-in-means between the two regions and its 95\% confidence
  interval.  Are sex ratios different across these regions?  What
  assumption is required in order for us to interpret this difference
  as causal?
  
## Answer 1

```{r}
## load data
women <- read.csv("data/chinawomen.csv", header = TRUE)
women$teadum <- ifelse(women$teasown > 0 & women$orch == 0, 1, 0)
women$orchdum <- ifelse(women$orch > 0 & women$teasown == 0, 1, 0)
women.after <- women[women$biryr == 1985, ]
## function to compute standard errors, confidence intervals
est.conf <- function(x, conf.level){
  mean.x <- mean(x, na.rm = TRUE)
  se.x <- sqrt(var(x) / length(x))
  ci <- c(mean.x - qnorm(1 - conf.level / 2) * se.x, 
          mean.x + qnorm(1 - conf.level / 2) * se.x)
  final <- c(mean.x, se.x, ci, length(x))
  names(final) <- c("mean", "se", "lower.ci", "upper.ci", "n.obs")
  return(final)
}
## point estimates and CIs
tea <- est.conf(women.after$sex[women.after$teadum == 1], 
                conf.level = 0.05)
orch <- est.conf(women.after$sex[women.after$orchdum == 1], 
                 conf.level = 0.05)
## difference-in-means
conf.level <- 0.05
est.diff <- tea[1] - orch[1]
se.diff <- sqrt(tea[2]^2 / tea[5] + orch[2]^2 / orch[5])
cis.diff <- c(est.diff - qnorm(1 - conf.level / 2) * se.diff, 
              est.diff + qnorm(1 - conf.level / 2) * se.diff)
results <- c(est.diff, se.diff, cis.diff)
names(results) <- c("diff", "se", "lower.ci", "upper.ci")
results
```


We find that the difference in sex ratios across the regions is
`r round(est.diff, 3)`, with a 95 percent confidence interval
equal to [`r round(cis.diff, 3)` ]. Since the interval contains
zero, it appears that sex ratios do not significantly differ across
the regions after the introduction of the reforms.

## Question 2

Repeat the analysis in the previous question for subsequent
  years, i.e., 1980, 1981, 1982, ..., 1990.  Create a graph which
  plots the difference-in-means estimates and their 95\% confidence
  intervals against years.  Give a substantive interpretation of the
  plot.
  
## Answer 2

```{r}
par(cex = 1.5) 
years <- unique(women$biryr[women$biryr > 1980])
diff <- matrix(data = NA, nrow = length(years), ncol = 5)
rownames(diff) <- years
colnames(diff) <- c("est", "se", "lower.ci", "upper.ci", "n.obs")
for(i in 1:length(years)){
  tea <- est.conf(women$sex[women$teadum == 1 & women$biryr == years[i]], 
                  conf.level = 0.05)
  orch <- est.conf(women$sex[women$orchdum == 1 & women$biryr == years[i]], 
                   conf.level = 0.05)
  conf.level <- 0.05
  est.diff <- tea[1] - orch[1]
  se.diff <- sqrt(tea[2]^2/tea[5] + orch[2]^2 / orch[5])
  cis.diff <- c(est.diff - qnorm(1 - conf.level/2) * se.diff, 
                est.diff + qnorm(1 - conf.level/2) * se.diff)
  diff[i, ] <- c(est.diff, se.diff, cis.diff, tea[5] + orch[5])
}
plot(years, diff[, "est"], type = "l",
     xlab = "Year", ylab = "Difference in Sex Ratio",
     ylim = c(min(diff[, "lower.ci"]) - 4 * sd(diff[, "lower.ci"]), 
              max(diff[, "upper.ci"]) + 4 * sd(diff[, "upper.ci"])), 
     main = "Post-Reform Sex Ratio Difference")
lines(years, diff[, "lower.ci"], lty = 2, col = "red")
lines(years, diff[, "upper.ci"], lty = 2, col = "red")
abline(h = 0, lty = 3)
```

The plot shows that despite some variation across the years, the
confidence intervals remain quite large.  As such, there does not
appear to be a change over time in our difference-in-means estimates
in the period after the reform was implemented. A slight dip occurs at
the beginning and end of the periods, however, which would correspond
to an increase in the sex ratio in fruit-producing counties relative
to tea-producing counties.

## Question 3

Next, we compare tea-producing and orchard-producing regions
  before the policy enactment.  Specifically, we examine the sex ratio
  and the proportion of Han Chinese in 1978. Estimate the mean
  difference, its standard error, and 95\% confidence intervals for
  each of these measures between the two regions. What do the results
  imply about the interpretation of the results given in Question~1?
  
## Answer 3

```{r}
women.before <- women[women$biryr == 1978, ]
han.tea <- est.conf(women.before$han[women.before$teadum == 1], 
                    conf.level = 0.05)
sex.tea <- est.conf(women.before$sex[women.before$teadum == 1], 
                    conf.level = 0.05) 
han.orch <- est.conf(women.before$han[women.before$orchdum == 1], 
                     conf.level = 0.05)
sex.orch <- est.conf(women.before$sex[women.before$orchdum == 1], 
                     conf.level = 0.05) 
## Han difference
han.diff <- han.tea[1] - han.orch[1]
han.se <- sqrt(han.tea[2]^2 / han.tea[5] + 
                   han.orch[2]^2 / han.orch[5])
han.cis <- c(han.diff - qnorm(1-conf.level / 2) * han.se, 
             han.diff + qnorm(1-conf.level / 2) * han.se)
results <- c(han.diff, han.se, han.cis)
names(results) <- c("mean", "se", "lower.ci", "upper.ci")
results
## sex difference
sex.diff <- sex.tea[1] - sex.orch[1]
sex.se <- sqrt(sex.tea[2]^2 / sex.tea[5] + 
                   sex.orch[2]^2 / sex.orch[5])
sex.cis <- c(sex.diff - qnorm(1 - conf.level / 2) * sex.se, 
             sex.diff + qnorm(1 - conf.level / 2) * sex.se)
results <- c(sex.diff, sex.se, sex.cis)
names(results) <- c("mean", "se", "lower.ci", "upper.ci")
results
```

We find a statistically significant difference in the Han Chinese proportion variable
between tea-producing and orchard-producing regions before the policy
enactment (but not the sex ratio variable), suggesting that these two regions may not be comparable.  This
casts a doubt on the validity of findings in Question 1.

## Question 4

Repeat the analysis for the sex ratio in the previous question
  for each year before the reform, i.e., from 1962 until 1978.  Create
  a graph which plots the difference-in-means estimates between the
  two regions and their 95\% confidence intervals against years.  Give
  a substantive interpretation of the plot.

## Answer 4

```{r}
par(cex = 1.5)
## sex differences
years <- unique(women$biryr[women$biryr < 1978])
results.sex <- matrix(data = NA, ncol = 4, nrow = length(years))
rownames(results.sex) <- years
colnames(results.sex) <- c("mean", "se", "lower.ci", "upper.ci")
for(i in 1:length(years)){
    women.before <- women[women$biryr == years[i], ]
    sex.tea <- est.conf(women.before$sex[women.before$teadum == 1], 
                        conf.level = 0.05)
    sex.orch <- est.conf(women.before$sex[women.before$orchdum == 1], 
                         conf.level = 0.05)
    ## sex difference
    sex.diff <- sex.tea[1] - sex.orch[1]
    sex.se <- sqrt(sex.tea[1]^2 / sex.tea[5] + 
                       sex.orch[1]^2 / sex.orch[5])
    sex.cis <- c(sex.diff - qnorm(1 - conf.level / 2) * sex.se, 
                 sex.diff + qnorm(1 - conf.level / 2) * sex.se)
    results.sex[i, ] <- c(sex.diff, sex.se, sex.cis)
}
## graph the results
plot(years, results.sex[, "mean"], type = "l",
     xlab = "Year", ylab = "Difference in Sex Ratio",
     ylim = c(min(results.sex[, "lower.ci"]) - 
                  4 * sd(results.sex[, "lower.ci"]), 
              max(results.sex[, "upper.ci"]) + 
                  4 * sd(results.sex[, "upper.ci"])), 
     main = "Pre-Reform Sex Ratio Difference")
lines(years, results.sex[, "lower.ci"], lty = 2, col = "red")
lines(years, results.sex[, "upper.ci"], lty = 2, col = "red")
abline(h = 0, lty = 3)
```

Similar to the other analyses, we see that while there is often a difference
in the sex ratio, it remains small and the confidence intervals for
each year are relatively large. There does not appear to be a
meaningful change in the difference between the sex ratios of the two
regions in the period leading up to the reforms.

## Question 5

We will adopt the difference-in-differences design by comparing
  the sex ratio in 1978 (right before the reform) with that in 1980
  (right after the reform).  Focus on a subset of counties that do not
  have missing observations in these two years.  Compute the
  difference-in-differences estimate and its 95\% confidence interval.
  Note that we assume independence across counties but account for
  possible dependence across years within each county.  Then, the
  variance of the difference-in-differences estimate is given by:
  \begin{eqnarray*}
    & & \V\{(\overline{Y}_{{\text tea}, {\text after}} -  \overline{Y}_{{\text tea},
    {\text before}}) - (\overline{Y}_{{\text orchard}, {\text after}} -  \overline{Y}_{{\text orchard},
    {\text before}})\} \\
    & = & \V(\overline{Y}_{{\text tea}, {\text after}} -  \overline{Y}_{{\text tea},
    {\text before}}) + \V(\overline{Y}_{{\text orchard}, {\text after}} -  \overline{Y}_{{\text orchard},
    {\text before}}) 
  \end{eqnarray*}
  where dependence across years is given by:
  \begin{eqnarray*}
     \V(\overline{Y}_{{\text tea}, {\text after}} -  \overline{Y}_{{\text tea},
    {\text before}}) 
    & = & \V(\overline{Y}_{{\text tea}, {\text after}}) - 2 {\rm
          Cov}(\overline{Y}_{{\text tea}, {\text after}}, \overline{Y}_{{\text tea},
          {\text before}}) + (\overline{Y}_{{\text tea}, {\text before}}) 
  \end{eqnarray*}
  
  A similar formula can be given for orchard-producing regions.  What
  substantive assumptions does the difference-in-differences design
  require?  Give a substantive interpretation of the results.

## Answer 5

```{r}
## Tea-producing regions
counties.before <- unique(women$admin[women$biryr == 1978])
counties.after <- unique(women$admin[women$biryr == 1980])
counties.list <- intersect(counties.before, counties.after)
women.before <- women[women$teadum == 1 & women$biryr == 1978 & 
                          women$admin %in% counties.list, ]
women.after <- women[women$teadum == 1 & women$biryr == 1980 & 
                         women$admin %in% counties.list, ]
women.before <- women.before[order(women.before$admin), ] # sort by admin
women.after <- women.after[order(women.after$admin), ]
tea.before <- est.conf(women.before$sex, conf.level = 0.05)
tea.after <- est.conf(women.after$sex, conf.level = 0.05)
covar.tea <- cov(women.before$sex,  women.after$sex, 
                 use = "pairwise.complete.obs") / nrow(women.before)
est.tea <- tea.after[1] - tea.before[1]
var.tea <- tea.after[2]^2 + tea.before[2]^2 - 2*covar.tea
## Fruit-producing regions
women.before <- women[women$orchdum == 1 & women$biryr == 1978 & 
                          women$admin %in% counties.list, ]
women.after <- women[women$orchdum == 1 & women$biryr == 1980 & 
                         women$admin %in% counties.list, ]
women.before <- women.before[order(women.before$admin), ] # sort by admin
women.after <- women.after[order(women.after$admin), ]
orch.before <- est.conf(women.before$sex, conf.level = 0.05)
orch.after <- est.conf(women.after$sex, conf.level = 0.05)
covar.orch <- cov(women.before$sex,  women.after$sex, 
                 use = "pairwise.complete.obs") / nrow(women.before)
est.orch <- orch.after[1] - orch.before[1]
var.orch <- orch.after[2]^2 + orch.before[2]^2 - 2*covar.orch
## Difference-in-differences
est.did <- est.tea - est.orch
se.did <- sqrt(var.tea + var.orch)
cis.did <- c(est.did - qnorm(1 - conf.level/2) * se.did, 
             est.did + qnorm(1 - conf.level/2) * se.did)
results <- c(est.did, se.did, cis.did)
names(results) <- c("diff-in-diff", "se", "lower.ci", "upper.ci")
results
```

Somewhat surprisingly, the difference-in-difference estimates suggest
that the tea-producing regions saw a relative increase of `r round(results[1], 3)` 
in the proportion of males born when
compared to the fruit-producing regions. However, the 95 percent confidence
interval for our estimate is [`r round(cis.did, 3)`], suggesting
that this difference is not statistically significant. The assumption we make is the
parallel time trend assumption, which requires that the over-time
change in the tea-producing regions would have been the same as in the
fruit-producing regions.
